\preparagraphspacing{}
\section*{File System Module Interfaces}
\label{sec:interfaces}

A complete KudOS configuration is composed of many modules.
By breaking file system code into
small, stackable modules, we are able to significantly increase code reuse.
We add an additional interface that helps to divide file system
implementations into common (reusable) code and file system-specific code.
We call this intermediate interface the ``Low-level File System'' (LFS). This
new interface is a departure from other stackable module systems,
like FiST~\cite{zadok00fist}, which stack higher-level operations.

The LFS interface has functions to allocate blocks, add blocks to
files, allocate file names, and other file system micro-ops. A module
implementing the LFS interface should define how bits are laid out on
the disk, but doesn't have to know how to combine the micro-ops
into larger, more familiar file system operations. 
A generic VFS-to-LFS module decomposes the larger file write,
read, append, and other standard operations into LFS micro-ops. This
one module can be used with many different LFS modules implementing
different file systems.

Figure~\ref{fig:kfs-graph} shows a contrived example taking advantage of the
LFS interface and change descriptors. A file system image is mounted
with an external journal, both of which are loop devices on the root file
system, which uses soft updates.
%
The journalled file system's ordering requirements are sent through the loop
device as change descriptors, allowing dependency information to be maintained
across boundaries that might otherwise lose that information.
%
In contrast, without change descriptors and the ability to forward change
descriptors through loop devices, BSD cannot express soft updates'
consistency requirements through loop-back file systems.
