\section {Evaluation}
\label{sec:evaluation}

We evaluate our prototype implementation of \Kudos\ in two ways: first, we
show that the overall performance is within reason, even though it is slower
than Linux or BSD by a nontrivial amount. We justify this difference, and argue
that it can be improved substantially. Second, we show specific block write
orderings in a variety of situations, demonstrating the effect and utility of
\opgroups\ and the potential for them to improve the efficiency of applications
using them.

\subsection {Overall Performance}

Our first benchmark test is to untar the Linux 2.6.15 source code, from
\texttt{linux-2.6.15.tar} which is already decompressed and cached in RAM. We
did this test on identical machines running Linux 2.6.12 (using ext3), FreeBSD
5.4 (using UFS, with and without soft updates), and \Kudos\ running as a Linux
2.6.12 kernel module (using UFS, with and without soft updates). As a second
test, we then delete the resulting source tree. The times below include time to
fully sync the changes to disk.

% make this into a table
\begin{figure}[htb]
\centering{
\begin{tabular}{|l|r|r|} \hline
System & Untar (sec) & Delete (sec) \\ \hline\hline
\Kudos\ (SU) & 24.50 & 11.69 \\ \hline
\Kudos\ (no SU) & & \\ \hline
FreeBSD 5.4 (SU) & 20.47 & 4.71 \\ \hline
FreeBSD 5.4 (no SU) & 17.74 & 15.14 \\ \hline\hline
% double line to separate linux
Linux 2.6.12 & 12.90 & 4.90 \\ \hline
\end{tabular}
}
\caption{\label{fig:macro} Untar/delete times for \Kudos, FreeBSD, and Linux,
with and without soft updates (SU) where possible.}
\end{figure}

\Kudos\ is about 20\% slower than FreeBSD using soft updates for
writes, and a little over a factor of two slower for deletions. This
shows that, although it is slow, \Kudos\ is close to running in
acceptable time. While we do have CPU usage and I/O delay problems, we
have only focused on performance for a few weeks, in which we were
able to speed up the system enormously.  There are more optimization
opportunities that can likely bring us very close in performance to
existing implementations of soft updates (see \S\ref{sec:discussion}).

\subsection {Block Writes}
We have also looked at the number of block writes that \Kudos\ makes
relative to other systems for several operations. In one test, we
create 100 small files in a directory and measure the number of blocks
written to disk. We do this with and without soft updates. \Kudos\
writes 122 blocks, while FreeBSD writes 135 blocks.

We also look at a very small benchmark: removing a file from a
directory and adding a new file to it. We compare the number of blocks
written by \Kudos\ with that of FreeBSD using soft updates. Here
\Kudos\ writes 15 blocks, while FreeBSD writes 83 blocks.

From this we gather that our UFS and soft updates implementations are
correct, and this reinforces our case that we need to reduce CPU
ussage to be competative.

\subsection {UW IMAP Case Study}
\label{sec:evaluation:uwimap}

FIXME: write about Figure~\ref{fig:imap-compare}.
Show fewer writes in UW IMAP using \opgroups\ (see \S\ref{sec:opgroup:uwimap})

\begin{figure}[htb]
\centering{
\begin{tabular}{|l|r|} \hline
System & \# writes (blocks) \\ \hline\hline
\Kudos\ (SU) & 1090 \\ \hline
%\Kudos\ (no SU) & & \\ \hline
FreeBSD 5.4 (SU) & 2340 \\ \hline
FreeBSD 5.4 (sync) & 3600 \\ \hline
\end{tabular}
}
\caption{\label{fig:imap-compare} Number of block writes to move 100
  IMAP messages.}
\end{figure}
