\section {Future Work}
\label{sec:future}

There are several areas in which we would like to expand our work. The obvious
first area we would like to work on is the performance of the KFS system. We
have already improved the performance by literally several orders of magnitude,
since we only recently began examining performance in addition to correctness,
but the system is not as fast as it could be -- and not quite as fast as it
needs to be to be a viable option for most computer systems.

There are several ways in which the performance can be improved. First, we
create a very large number of change descriptors, and the sheer number of them
can cause problems for any of our algorithms which needs to traverse parts of
the change descriptor dependency graph. We can attack this problem from two
sides: we can reduce the number of change descriptors by intelligently merging
them when we determine that having separate change descriptors is not necessary,
or we can improve the efficiency of the traversal algorithms to need to examine
fewer change descriptors. We have worked on both of these approaches, especially
the second, but we believe further improvements can be made.

Another item we would like to add is an ext2 file system component. Currently,
we have written only a UFS component, aside from a component for a very simple
prototype file system called JOSFS. Once we have an ext2 component, we could use
our journal module to get ``ext3b''. Implementing an ext2 component would also
help us to identify more opportunities for sharing code between file system
implementations, and ways in which our interfaces might accomodate different
file systems more neutrally.

\begin{itemize}
\item Fast cycle checking
\item Better cache eviction algorithm
\item Transactional user-level change descriptors
\item More flexible rules for user-level change descriptors
\item Metadata-only journaling
\item More references
\end{itemize}
