\section{Consistency}
\label{sec:consistency}

To better understand factors that influenced our design decisions, we examine
file system consistency protocols and shortcomings in their implementations.
Two mainstream methods for file system consistency are journaling and soft
updates. Journaling has been around longer and this is evident by the number of
file systems that use it: ext3, HFS+, JFS, NTFS, Reiserfs, and XFS, to name a
few. In contrast, soft updates is a relatively new concept and is only widely
used for UFS on the BSD family of operating systems.

\input{journaling}

\input{softupdates}

\subsection{Current Implementations}
\label{sec:consistency:implementations}
While implementations of journaling and soft updates protect file systems
and provide good performance, they leave much to be desired in terms of
modularity, code reuse, and simplicity. Many file system implementations have
journaling code that is closely-coupled with the rest of the file system.
For instance, all the journaled file systems on Linux have their own code for
journaling. Although there have been attempts to provide journaling capability
as a separate \module, none of the attempts have been successful. On FreeBSD,
the Gjournal project for GEOM is incomplete and cannot replace a proper
journaled file system. Apple's OS X provides a generic journaling layer, but
thus far it has only been used with HFS+. Similarly, Linux's ext3
implementation is the sole user of Linux's generic journaling block device.
The source code for ext3 also has much in common with the code for ext2, its
non-journaled predecessor, yet there is little code sharing between the two
other than common header files. For soft updates, the fundamental idea of
sequencing writes is built into the implementation, thus making modularity
hard or impossible. The soft updates implementations can improve on
simplicity, however. For UFS on FreeBSD, soft updates uses 16 different types
of data structures to manage dependencies for various parts of the file
system. Though there are advantages to having specialized dependency
structures, it is more difficult to understand them and manage their
interactions with each other.

