\subsection{Soft Updates}
\label{sec:softupdates}

Soft updates takes a different approach where the system carefully orders
writes to disk. By following a few safety principles, soft
updates' \emph{update dependencies} ensures the disk is consistent after every
write. This gives file system integrity guarantees without a journal and the
associated redundant writes. 

For soft updates, the three golden rules are: never point to a structure
before it has been initialized; never reuse a resource before nullifying
all previous pointers to it; and never reset the last pointer to a live
resource before a new pointer has been set. Although it seems easy to just
follow these rules and write changes out to disk in a proper order, the
reality of the situation is not that simple. Caches flush data out to disk
in the unit of sectors or blocks, not in terms of byte-level changes. This
leads to situations where the cache needs to write changes out to disk, but
the byte-level dependencies create cyclic dependencies at the block level.
To solve this problem, soft updates uses \emph{dependency structures} to
retain the state of data structures before and after changes occur. With the
ability to undo changes, soft updates can break block-level cyclic dependencies
by reverting changes that cannot be safely written to disk. Once the cycle has
been broken, soft updates restores the reverted changes and writes them out at
a later time.
%Compared to an unsafe file system, this presents a small
%overhead, as it does generate more writes. With respect to journaling, soft
%updates writes less to the disk, since it does its bookkeeping in memory.
%
%There are some drawbacks to soft updates for consistency and crash recovery.
%Whereas a journaled file system can refer to the journal, soft updates has no
%persistent information available after a crash. The soft updates rules protect
%against inconsistencies that leads to corrupt data, but it does not protect
%against resource leaks. As a result, unused disk blocks and inodes can be
%marked as used, and inodes may have higher link counts than actual link counts.
%To correct these problems, a file system using soft updates still requires
%a file system check, though it can be done in the background while the file
%system is online.

