% -*- mode: latex; tex-main-file: "paper.tex" -*-

\begin{abstract}

%% The reliability of a file system to correctly store and later provide
%%  access to data is one of its most important properties.
%% %
%% File systems today deal with many challenges that make implementing this
%%  reliability difficult: power losses, software failures, and even user
%%  intervention all pose significant threats.
%% %
%% In order to avoid time-consuming and potentially ineffective manual
%%  checks like \emph{fsck} when recovering from failures, file systems use
%%  a variety of techniques.
%
Reliable storage systems depend at least in part on a ``write-before''
 relationship, where one change to stable storage is delayed until
 another change commits.
%
A journaled file system, for example, must write and commit each
 journal transaction before applying that transaction's changes, and
 soft updates~\cite{ganger00soft} and other consistency enforcement
 mechanisms have similar constraints, implemented in each case in
 system-dependent ways.
%
%% These relationships are implemented in system-dependent ways.
%
We present a general \emph{\patch} abstraction that makes the write-before
 relationship
 explicit and file system agnostic.
%
A \patch-based file system implementation expresses dependencies among
 writes to stable storage, but does not enforce specific block write orders
 to satisfy those dependencies.
%
Storage system \modules\ can examine, preserve, and modify the dependency
 structure; the buffer cache writes blocks as constrained by that structure.
%
These generalized file system dependencies are naturally exportable to
 user level.
%
Our patch-based storage system, \emph{\Featherstitch}, includes a number of
 important optimizations.
%
A \Kudos\ ext2 prototype running in the Linux kernel supports asynchronous
 writes, soft updates-like dependencies, and journaling.
%
It performs competitively with ext2 and ext3 on many benchmarks;
 supports unusual configurations, such as correct dependency
 enforcement within a loopback file system; and allows applications
 to specify consistency requirements without micromanaging how those
 requirements are satisfied.


\begin{comment}

File systems ensure that their data is kept consistent through careful
 write ordering, where certain disk blocks must be committed to stable
 storage before other blocks.
%
Previous file systems have enforced write orderings in system-dependent
 ways, either with rules specialized for each file system
 structure~\cite{ganger00soft} or with a journal, which enforces a
 particular consistency protocol.
%
We present a general \emph{\patch} abstraction that can represent any
 write ordering in a file system agnostic manner.
%
A \patch-based file system implementation expresses dependencies among
 writes, but does not enforce specific block write orders that satisfy
 those dependencies.
%
Storage system \modules\ can examine, preserve, and modify write
 orderings.
%
Generalized file system dependencies are naturally exportable to user
 level, allowing applications to specify their own consistency protocols
 for the storage system to follow.

We present the \patch\ abstraction, describe a number of important
 optimizations for \patch-based storage systems, and present a Linux kernel
 implementation of a storage subsystem that uses \patches\ to enforce
 consistency.
%
Our ext2 prototype is competitive with Linux ext2 and ext3 and allows
 several novel configurations, such as ext2 with soft updates or correct
 dependency enforcement within a loopback file system, and provides a
 simple interface for user applications to directly affect \patches.

\end{comment}


\begin{comment}

We propose a file system implementation architecture, called \emph{\Kudos},
where structures called \emph{\patches} represent any and all changes to
stable storage.
%
%%  File systems generate \patches\ for all writes, then
%% send them to block devices for eventual commit. Explicit dependencies between
%% \patches\ let \Kudos\ \modules\ preserve necessary file system
%% invariants without understanding the file system itself. \Patches\ can
%% implement many consistency mechanisms, including soft updates and journaling.
%
\Kudos\ is decomposed into fine-grained \modules\ which generate, consume,
 forward, and manipulate \patches.
%
The uniform abstraction of \patches\ allows modules to impose and
 follow arbitrary file system consistency policies: a collection of
 loosely-coupled modules cooperates to implement strong and possibly
 complex guarantees, even though each individual module does a relatively
 small part of the work.
%
%% A particular innovation of the
%% \module\ design is the separation of the low-level specification of on-disk
%% layout from higher-level file system-independent code, which operates on
%% abstract disk structures. 
%
For example, by observing and modifying \patch\ constraints, our
 journaling \module\ can automatically add journaling to any file system.
%
Additionally, a new system call interface gives applications some direct
 control over \patches. We have used this interface to
improve the UW IMAP server, removing inefficient and unnecessary calls to
\texttt{fsync()} while preserving the integrity of mail messages.
%
We have implemented \Kudos\ as a Linux kernel module. Our current
implementation is competitive with FreeBSD soft updates for number of
blocks written, and allows several novel configurations like ext2 with
soft updates or correct UFS soft updates over a loopback device.

\end{comment}

\end{abstract}
