% -*- mode: latex; tex-main-file: "paper.tex" -*-

\section {Introduction}
\label{sec:intro}

\begin{comment}
This paper aims to evaluate whether a simple, unified abstraction that
represents all modifications to stable storage, including
\emph{dependencies} among modifications, can be used to efficiently
implement a complete file system layer, where modifications are common
and cache sizes are large.
%
The answer is a qualified yes.
\end{comment}


This paper builds a complete and relatively efficient file system layer,
 called \emph{\Featherstitch}, that uses
 a simple, uniform, and file system agnostic data type to explicitly
 represent all ``write before'' relationships among changes to stable
 storage.
%
``Write before'' relationships, which require some changes to be committed
 before others, are part of the abstract basis of every mechanism for ensuring file system
 consistency and reliability, from journaling to synchronous writes.
%
\Kudos\ API design and optimizations make these relationships, in a concrete form
 called \emph{patches}, a promising implementation strategy as well.


\begin{comment}
As file system functionality increases, maintaining file system
 correctness in the presence of failures is increasingly a focus of
 research~\cite{sivathanuetal05-logic,denehyetal05-journal-guided}.
%
File systems today deal with many challenges that make implementing this
 property difficult: power losses, software failures, and even user
 intervention all pose significant threats.
%
To meet this challenge, file systems use a variety of techniques, like
 journaling and soft updates.
%
These mechanisms are each based on imposing some ``write before''
 relationship among buffered changes to the data in stable storage.
%
The answer is a qualified yes.
\end{comment}


A patch represents both a change to disk data and any \emph{dependencies}
 between that change and other changes. 
%
\Patches\ were initially inspired by the dependency abstraction from BSD's soft
 updates~\cite{ganger00soft}, but whereas soft updates dependencies
 implement a particular type of consistency
 and involve many structures specific to the UFS file
 system~\cite{mckusick99soft}, \patches\ are fully general,
 specifying only how a range of bytes should be changed.
%
This lets file system implementations specify a
 write-before relationship between changes without dictating
% (or worrying about)
 a write order that honors that relationship.
%
It lets storage system components examine and
 modify dependency structures independent of the file system's layout,
 possibly even changing one type of consistency into another.
%
It also lets applications modify dependency structures: a
 user-level interface called \emph{\patchgroups} shows that applications
 can access \patches\ and define consistency policies for the underlying
 storage system to follow.


A uniform and pervasive \patch\ abstraction may simplify the
 implementation of, extension of, and experimentation with mechanisms for
 ensuring file system consistency and reliability.
%
File system implementers find it difficult to provide
 consistency guarantees~\cite{tweedie98journaling,mckusick99soft}
 and implementations are often buggy~\cite{yang04using,yang06explode},
 a situation further complicated by file system extensions and
 special disk interfaces~\cite{soules03metadata,fast04versionfs,quinlan02venti,cornell04wayback,wright03ncryptfs,sivathanu03semantically-smart,sivathanu05database-aware}.
%
File system extension techniques such as stackable file
 systems~\cite{zadok00fist,zadok99extending,heidemann94filesystem,rosenthal90evolving}
 leave consistency up to the underlying file system; any extension-specific
 ordering requirements are difficult to express at the VFS layer.
%
Although maintaining file system
 correctness in the presence of failures is increasingly a focus of
 research~\cite{sivathanuetal05-logic,denehyetal05-journal-guided},
%
other proposed systems for improving file system integrity
 differ mainly in the kind of consistency they aim to impose, ranging from
 metadata consistency to full data journaling and full ACID
 transactions~\cite{gal05transactional,liskov04transactional}.
%
Some users, however, provide end-to-end reliability for some data
 and prefer to avoid \emph{any} associated consistency
 slowdowns in the file system layer~\cite{googleext2}.
%
\Patches\ can represent all these choices, and since they provide a common
 language for file systems and extensions to discuss consistency
 requirements, they can even allow combinations of consistency protocols to
 comfortably coexist.


\begin{comment}
But different extensions within a file system, or different
 applications over the file system, may require different types of
 consistency semantics, and performance suffers when lower layers are
 unnecessarily denied the opportunity to reorder writes;
 %
\Patches\ can implement many consistency mechanisms, including
 soft updates and journaling, and can allow combinations of different
 consistency protocols to exist at the same time.
 %
\Patches\ provide a simple and effective way for such extensions to
 express their requirements of the storage system, adding to the
 requirements already expressed by the file system itself.
\end{comment}


Applications likewise have few mechanisms
 for controlling buffer cache behavior in today's systems, and
%
robust applications, including databases, mail servers, and source code
 management tools, must choose between several mediocre options.
%
They can accept the performance penalty of expensive system calls like
 \texttt{fsync} and \texttt{sync}, use tedious and fragile sequences
 of operations that count on particular file system semantics, or
 rely on specialized raw-disk interfaces.
%
\Patchgroups\ give applications some of the same benefits that \patches\ provide
 for kernel-level file system implementations and extensions.
%
Modifying an IMAP mail server to use \patchgroups\ required only localized
 changes.  The result meets IMAP's consistency requirements no
 matter what type of consistency the underlying file system provides, and
 avoids the performance hit of full synchronization.


\begin{comment}
Our file system modules impose soft updates-style \patch\ requirements by
 default, since doing so requires some knowledge of the file system's
 structures; we have also written a journal module that can change
 existing dependencies to express either full or metadata-only journaling.
%
A file system module not interested in supporting soft updates support
 could instead impose no \patch\ requirements, and count on the journal
 module to provide a consistency guarantee.


The \Kudos\ storage system implementation is decomposed entirely into
 pluggable \modules\ that manipulate \patches, hopefully making the system
 as a whole more configurable, extensible, and easier to understand.
%
Any storage system \module\ can generate \patches; other modules can examine
 them and modify them when required.
%
\Patch\ dependencies are obeyed by all other storage system layers, allowing
 them to be passed through layers such as loopback block devices.
%
As a result, the loosely-coupled \modules\ that implement a file system
 can cooperate to enforce strong and often complex consistency guarantees,
 even though each \module\ only does a small part of the work.
\end{comment}


Though production file systems use system-specific optimizations to achieve
 consistency without sacrificing performance, we tried to achieve good
 performance in a general way.
%
A naive \patch-based implementation scaled terribly,
 spending far more space and time on dependency manipulation than
 conventional systems.
%
We therefore developed several optimizations that attack \patch\ memory and
 CPU overheads, reducing some overheads by at least \patchoptundo\ in our
 tests. 
%
Room for improvement remains, but \Featherstitch\ is
 competitive with equivalent Linux configurations on many benchmarks.


Our contributions include the \patch\ model and design, our
 optimizations for making \patches\ more efficient,
 the \patchgroup\ mechanism that exports
 \patches\ to applications, and several individual \Kudos\ modules, such as
 the journal.


We describe \patches\ abstractly, state their behavior and safety
 properties, give examples of their use, and reason about the
 correctness of our optimizations.
%
We then describe the \Kudos\ implementation, which is decomposed
 entirely into pluggable modules that manipulate
 \patches, hopefully making the system as a whole more configurable,
 extensible, and easier to understand.
%
We conclude following an evaluation comparing \Kudos\ and Linux-native file
 system implementations.




\begin{comment}
%
Our benchmarks show that our optimizations can reduce the number of
 \patches\ \Kudos\ creates by \patchoptcount\ and the amount of undo data
 memory it allocates by \patchoptundo.
%
Our prototype is not yet as fast as we would like, but it is competitive
 with Linux on many of our benchmarks.
\end{comment}