% -*- mode: latex; tex-main-file: "paper.tex" -*-

\section{\Patch\ Model}
\label{sec:patch}

\makeatletter
\let\emptyset\varnothing
\newcommand{\PState}[1]{\ensuremath{#1.\textit{state}}}
\newcommand{\PBlock}[1]{\ensuremath{#1.\textit{block}}}
\newcommand{\PMemst}{\ensuremath{\textit{mem}}}
\newcommand{\PInfst}{\ensuremath{\textit{flight}}}
\newcommand{\PDiskst}{\ensuremath{\textit{disk}}}
\newcommand{\PSetlim}[1]{\def\@next{#1}\ifx\@next\@empty\else[\@next]\fi}
\newcommand{\PMem}[1][]{\ensuremath{\textit{Mem}\PSetlim{#1}}}
\newcommand{\PInf}[1][]{\ensuremath{\textit{Flight}\PSetlim{#1}}}
\newcommand{\PDisk}[1][]{\ensuremath{\textit{Disk}\PSetlim{#1}}}
\newcommand{\PHard}[1][]{\ensuremath{\textit{\Nrb}\PSetlim{#1}}}
\newcommand{\PSoft}[1][]{\ensuremath{\textit{\Rb}\PSetlim{#1}}}
\newcommand{\PEmpty}[1][]{\ensuremath{\textit{\Noop}\PSetlim{#1}}}
\newcommand{\PDDepset}[1]{\ensuremath{\def\@next{#1}\ifx\@next\@empty\else\@next.\fi\textit{ddeps}}}
\newcommand{\PDepend}{\ensuremath{\leadsto}}
\newcommand{\PDDepend}{\ensuremath{\rightarrow}}
\newcommand{\PDepset}[1]{\ensuremath{\textit{Dep}[#1]}}
\newcommand{\PRDepset}[1]{\ensuremath{\textit{RDep}[#1]}}
\makeatother

Every change to stable storage in a \Kudos\ system is represented by a
\emph{\patch}.
%
This section describes the basic \patch\ abstraction using a semi-formal
notation that will be useful for analyzing our optimizations later.
%
Although the \patch\ idea would apply to any stable medium, to network file
systems, or to multiple disks, we use terms like ``the disk'' and ``the
disk controller'' throughout to simplify our terminology.


Each \patch\ $p$ encapsulates four important pieces of information: its
 \emph{block}, its \emph{state}, a set of \emph{direct dependencies}, and
 some \emph{undo data}.

\Patch\ $p$'s \textbf{block}, written $\PBlock{p}$, is the unit of disk data
 to which $p$ applies.  The disk controller is assumed to execute writes one
 block at a time (although not necessarily atomically; see below).  A file system
 change that affects $n$ blocks must be represented using at least $n$
 \patches, since a \patch\ by definition can touch only one block.

Each \patch\ is in one of three \textbf{states}: \emph{in memory}; \emph{in
 flight} to the disk controller, but not yet committed to disk; or
 \emph{on disk}.  The intermediate in-flight state is necessary because
 operating system software loses control of a \patch\ upon writing its block
 to disk.  $p$'s state is written $\PState{p} \in \{\PMemst, \PInfst,
 \PDiskst\}$.  The operating system changes a \patch's state from
 \PMemst\ to \PInfst\ by writing its block to the disk controller.  Some
 time later---after possible controller-level disk scheduling, transit over
 the system bus, and a period in the disk's cache---the disk writes the block
 to stable storage and reports success.  When the processor receives this
 notification, it changes the \patch's state to \PDiskst.
%
The sets \PMem, \PInf, and \PDisk\ are defined to contain all \patches\ with
 the given state, and the sets $\PMem[b]$, $\PInf[b]$, and $\PDisk[b]$ are
 defined to contain all \patches\ with the given state \emph{on the given
 block} $b$.

\Patch\ $p$'s \textbf{direct dependencies}, written $\PDDepset{p}$, is a set of
 other \patches\ on which $p$ ``depends''.
%
That is, every \patch\ in $\PDDepset{p}$ should be committed to disk either
 before, or atomically with $p$ (i.e. $p$ and $q$ are on the same block).
%
Dependencies are set by the file system based on the consistency
 semantics it wants to ensure.
%
For example, a file system with asynchronous writes (and no durability
 guarantees) might write all \patches\ with $\PDDepset{p} = \emptyset$;
%
a file system wishing to impose strictly ordered writes might set
 $\PDDepset{p_n} = \{p_{n-1}\}$.
%
In summary, an upper file system layer defines an initial set of
 dependencies;
%
the rest of \Kudos\ generally preserves this initial set, but can modify
 it as necessary to change dependency semantics;
%
and finally, the buffer cache obeys the constraints thus defined.


We say $p$ \emph{directly depends on} $q$, and write $p \PDDepend q$, when
 $q \in \PDDepset{p}$.
%
We say $p$ \emph{depends on} $q$, and write $p \PDepend q$, when there
 exists a dependency chain from $p$ to $q$: that is, either $p \PDDepend q$
 or, for some \patch\ $x$, $p \PDepend x \PDDepend q$.
%
\Patch\ $p$'s set of \emph{dependencies} is written $\PDepset{p} = \{ q \mid
 p \PDepend q \}$; this is of course a superset of its direct dependencies.
%
Given a \emph{set} of \patches\ $P$, we write $\PDepset{P}$ to mean the set's
 combined dependencies $\bigcup_{p\in P} \PDepset{p}$. Our \patch\ notation
 is summarized in Figure~\ref{fig:patchnot}.

\textbf{Example.}
%
Patches implementing a journal transaction could be arranged as follows:
%
the journal commit record $c$ would have $c \PDDepend \{j_{1\dots n}\}$,
where each $j_i$ writes an entry into the journal.
%
Then the file system updates described in the journal would each have
$u_i \PDDepend c$.
%
This arrangement of dependencies ensures that the commit record is not written
to disk until all the journal data is present there, and that the updates to the
file system itself are not written until the commit record is present on disk.


\textbf{Undo data.}
%
When a \patch\ is created, the buffer cache's copy of the block data
is modified in-place to reflect the specified change. However,
%
some arrangements of \patches\ may require that the buffer cache
\emph{not} write one or more \patches\ on some block.
%
In particular, a series of file system operations may create dependencies
that enforce a circular order among blocks, even though the dependencies
themselves do not form a cycle~\cite{ganger00soft}.
%
For instance, a \patch\ writing the directory entry for a new file \texttt{f1}
may depend on a \patch\ allocating an inode for it, and a \patch\ deleting the
entry \texttt{f2} from the same directory may be depended upon by a \patch\
freeing the inode for \texttt{f2}.
%
If the two inodes are stored in the same block, and the two directory entries
are in the same block as well, each block will have a \patch\ depending on the
other block.
%
This is problematic since blocks with circular dependencies can never be
written: no block can be written first since each block depends on another.

For this reason, each \patch\ carries \emph{undo} information that gives
the previous version of the range of block data altered by the \patch.
%
If a \patch\ $p$ is not written with its containing block, the system reverts
the \patch, which swaps the new data and the previous version on the (single)
copy of the block in the buffer cache.
%
Once the block is written, the system will re-apply the \patch\ and, when
allowed, write the block again, this time including the \patch.
%
Undo information adds greatly to memory and CPU utilization, but it can
often be optimized away, as we show below.

\begin{figure}[t]
\begin{small}
\begin{tabular}{@{}ll@{}}
$\PBlock{p} = b$ & $p$'s block \\
$\PState{p}$ & $p$'s state, $\in \{\PMemst, \PInfst, \PDiskst\}$ \\
$\PDDepset{p}$ & $p$'s direct dependencies: \patches\ that must go to disk \\
	& before $p$ \\
\noalign{\vskip3pt}
$\PMem$ & all in-memory \patches: $\{p \mid \PState{p} = \PMemst\}$ \\
$\PMem[b]$ & $\{p \mid \PState{p} = \PMemst \text{ and } \PBlock{p} = b
 \}$ \\
$\PInf, \PDisk$ & similarly for in-flight and on-disk \patches\ \\
\noalign{\vskip3pt}
$p \PDDepend q$ & $p$ directly depends on $q$: $q \in \PDDepset{p}$ \\
$p \PDepend q$ & $p$ depends on $q$: either $p \PDDepend q$ or there exists \\
       & a \patch\ $x$ so that $p \PDepend x \PDDepend q$ \\
$\PDepset{p}$ & $p$'s dependencies: $\{ q \mid p \PDepend q \}$ \\
\end{tabular}
\end{small}

\caption{\Patch\ notation.}
\label{fig:patchnot}
\end{figure}


\subsection{Obeying Dependencies}
\label{sec:patch:dependencies}

\Kudos\ must ensure the \textbf{disk safety property}, which states that
 the dependencies of all \patches\ on disk are also on disk:
%
\[ \PDepset{\PDisk} \subseteq \PDisk. \]
%
Thus, no matter when the system crashes, the disk is consistent in terms of
dependencies.
%
The file system's job is to set up dependencies so that the disk safety
property implies file system correctness.

However, \Kudos\ can only control when \patches\ are handed to the disk
 controller, not when they are written to disk.
%
Disk controller behavior is encapsulated in the following atomic action:

\begin{tabbing}
\textit{Commit block:} \\
\quad Pick some block $b$ with $\PInf[b] \neq \emptyset$. \\
\quad For each $p \in \PInf[b]$, set $\PState{p} \gets \PDiskst$.
\end{tabbing}

\noindent
%
Since the controller can write blocks in any order, \Kudos\ must ensure
that in-flight blocks are mutually independent.  This is precisely stated
by the \textbf{in-flight safety property:}
%
\[ \text{For all blocks $b$,~~} \PDepset{\PInf[b]} \subseteq \PInf[b] \cup \PDisk . \]
%
This ensures that for any $b' \neq b$, $\PDepset{\PInf[b]} \cap
 \PDepset{\PInf[b']} \subseteq \PDisk$; the disk controller can safely
 write in-flight blocks in any order and still preserve disk safety.
%
(The union with $\PInf[b]$ is necessary for the case of multiple in-flight
\patches\ on the same block.)


The buffer cache must thus behave according to the following atomic action:

\begin{tabbing}
\textit{Write block:} \\
\quad Pick some block $b$ with $\PMem[b] \neq \emptyset$ and $\PInf[b] =
\emptyset$. \\
\quad Pick some $P \subseteq \PMem[b]$ with $\PDepset{P} \subseteq P \cup
\PDisk$. \\
\quad For each $p \in P$, set $\PState{p} \gets \PInfst$.
\end{tabbing}

\noindent
%
We require that $\PInf[b] = \emptyset$ since the disk controller can write
in-flight blocks in any order, and thus at most one version of a block can
be in flight at any time.
%
The buffer cache is also expected to write all blocks eventually, a
 liveness property.

\textbf{Example.}
%
Suppose the buffer cache has \patches\ arranged as in the previous journal
example, and that $\{j_i\} \subseteq \PDisk$, i.e., the journal entries have
already been written to disk.
%
The buffer cache cannot write block $\PBlock{u_i}$ for any $i$, since $u_i
\PDDepend c$ and $\PState{c} \neq \PDiskst$.
%
Its only option is to write $\PBlock{c}$ since $\PDepset{c} = \{j_i\}
\subseteq \PDisk$.
%
So, via the action above, it sets $\PState{c} \gets \PDiskst$.
%
Later, when notification is received that $\PBlock{c}$ has been written to disk,
$\PState{c}$ is set to $\PDiskst$ and the buffer cache then has the option to
write any of $\PBlock{u_i}$.

This model does not completely define the disk's behavior on system crash,
 in particular with respect to in-flight blocks.
%
Soft updates inherently assumes that blocks are written
\emph{atomically}, except in the case of catastrophic media error:
%
if the disk fails while a block $b$ is in flight, then $b$'s
value on recovery must equal either the old value or the new value.
%
Most journal designs do not rely on this assumption, and can recover
 properly even if in-flight blocks are corrupted---for instance,
 because the memory holding the new value of the block lost its coherence
 before the disk stopped writing~\cite{nightingale06rethink}.
%
However, some disks may actually provide an atomicity guarantee, for
 instance by using non-volatile memory to store blocks before they make it
 onto disk.
%
The \Kudos\ core makes no assumptions about block atomicity, instead relying
 on software above it to implement a consistency protocol that makes sense
 for the given disk.


The buffer cache can't handle cross-block \patch\ dependency cycles since
it must write data in units of blocks.
%
Thus, if $p \PDepend q \PDepend p$, then $p$ and $q$ must be on the same
block; otherwise, neither $p$ nor $q$ could ever get written.
%
Whenever upper layers might generate cross-block dependency cycles,
intervening layers must break those cycles somehow before they reach the
cache.
%
For instance, a journal layer might detect dependency cycles and put the
corresponding \patches\ into a single transaction, but the
\emph{implementation} of that transaction could not contain cross-block
cycles.
%
In practice, the \Kudos\ implementation currently disallows any dependency
cycles, including cycles within a single block; this simplifies many graph
traversals used to determine when optimizations may be applied, and has not
yet caused any difficulty in implementing different consistency policies.


\input{noop}
