\section{Consistency Models}
\label{sec:using}

Soft updates, journaling, and other consistency models
correspond to different \patch\ arrangements.
%
\Kudos's current file system \modules\ generate dependencies corresponding
to soft updates orderings by default~\cite{ganger00soft}.
%
Adding an independent journal \module\ rearranges the \patches\ to provide
journaling semantics.

\paragraph{Soft updates}
\label{sec:using:softupdate}

In the BSD UFS implementation of soft updates, each UFS operation is
represented in memory by a structure encapsulating the disk changes
necessary to implement that operation. As a result, many
specialized data structures represent the different possible file system
operations. These structures, their relationships to one another, and their uses
for tracking and enforcing dependencies are quite
complex~\cite{mckusick99soft}.

\Kudos, in contrast, represents all dependencies by \patches\ and relies on
general optimizations to reduce memory usage.
%
The resulting system is not as optimized as the BSD implementation, but
much simpler to specify and easier for other modules to manipulate.
%
The \patches\ that make up a file system operation are connected to specify
the order in which the changes must be written to disk.
%
For instance, most file systems require at least two \patches\ to remove
a block from a file, one ($p$) to clear the block reference from the file's
block list and one ($q$) to mark the block as free; adding the $q \PDDepend
p$ dependency straightforwardly implements soft updates semantics.
%
Another example is given in Section~\ref{sec:patch:examples}.

\input{journal}

\paragraph{Asynchronous writes}
\label{sec:modules:unlink}

Finally, we also wrote a trivial module that removes all dependencies from
incoming \patches, allowing the buffer cache to write blocks in any order.
%
This implements similar semantics to existing file systems like ext2 in
asynchronous write mode.
